---
title: "OABS Accuracy Analysis"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
    theme: spacelab
    includes:
      after_body: footer.html
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

```{css, echo=FALSE}
.btn {
    border-width: 0 0px 0px 0px;
    font-weight: normal;
    text-transform: ;
}
.btn-default {
    color: #2ecc71;
    background-color: #ffffff;
    border-color: #ffffff;
}
```

```{r read in data, warning = FALSE, message=FALSE}

# GET LIBS AND READ IN DATA

library(tidyverse)
library(patchwork)
library(plotrix)
library(summarytools)
library(DT)
library(afex)
library(ez)
library(emmeans)

AccDat = read.csv('OABS_Accuracy_Data_deID.csv', header = TRUE)
names(AccDat)[1] = 'idonset'
AccDat = AccDat %>% filter(!is.na(condtria))

#Coerce all the MacSHAPA timing variables to times
MacSHAPA_format <- function (MacSHAPA_Col) {
  #browser()
  MacSHAPA_split = as.data.frame(MacSHAPA_Col)
  MacSHAPA_split = separate(MacSHAPA_split, MacSHAPA_Col, into = c("Hours", "Minutes", "Seconds", "Frames"), sep = ":")
  Cols_to_numConv = c('Hours', 'Minutes', 'Seconds', 'Frames')
  MacSHAPA_split = lapply(MacSHAPA_split, as.numeric)
  MacSHAPA_split$Milliseconds = MacSHAPA_split$Frames * 1000/60 # mult by converstion factor of 1000 ms / 60 fps *** Need to adjust fps when Simone provides actual number ***  ***Simone says frame rate 30 on 3-26-20, but it looks to me like 60 fps because the time stamps show the final digit >30....I JUST DON'T KNOW.  60: Final answer.
  MacSHAPA_split = lapply(MacSHAPA_split, as.character)
  MacSHAPA_split = as.data.frame(MacSHAPA_split)
  MacSHAPA_split = MacSHAPA_split %>%
    mutate_all(as.character) %>%
    unite(Seconds, Seconds, Milliseconds, sep = ".")
  #browser()
  MacSHAPA_fixed = unite(MacSHAPA_split, Time, Hours, Minutes, Seconds, sep = ":")
  MacSHAPA_fixed$Frames = NULL
  return(MacSHAPA_fixed)
}

for (i in grep("idonset", colnames(AccDat)):grep("stepoffset", colnames(AccDat))) {
  AccDat[,i] = MacSHAPA_format(AccDat[,i])
}

Cols_to_timeConv = colnames(AccDat[grep("idonset", colnames(AccDat)):grep("stepoffset", colnames(AccDat))])
AccDat[Cols_to_timeConv] = lapply(AccDat[Cols_to_timeConv], strptime, '%H:%M:%OS')
AccDat[Cols_to_timeConv] = lapply(AccDat[Cols_to_timeConv], as.POSIXct)

# # was working on calculating age....it was already in there as "testage"
# Cols_to_dateConv = colnames(AccDat[grep("tday", colnames(AccDat)):grep("bday", colnames(AccDat))])
# AccDat[Cols_to_dateConv] = lapply(AccDat[Cols_to_dateConv], as.character)
# AccDat[Cols_to_dateConv] = lapply(AccDat[Cols_to_dateConv], as.Date, '%m/%d/%Y')

#isolate SubNum
AccDat$SubNum = lapply(AccDat$subj, function(subj) {
  subj = as.character(subj)
  if (nchar(subj) == 4) {
    SubNum = substr(subj, 2, 2)
  } else if (nchar(subj) == 5) {
    SubNum = substr(subj, 2, 3)
  }
})
AccDat$SubNum = as.character(AccDat$SubNum)
AccDat = filter(AccDat, SubNum != "k8")
AccDat$SubNum = as.factor(AccDat$SubNum)

#Create step variable
AccDat$StepNum = 1
for (i in 1:nrow(AccDat)) {
  if (i != 1) {
    if (AccDat$condtria[i] == AccDat$condtria[i-1]) {
      AccDat$StepNum[i] = AccDat$StepNum[i-1] + 1
    } else {
      AccDat$StepNum[i] = 1
    }
  }
}

# Find instantaneous cadence (i.e., time elapsed between consecutive steps)
AccDat$IntStepInterval = NA
for (i in 1:nrow(AccDat)) {
  if (AccDat$StepNum[i] != 1) {
    AccDat$IntStepInterval[i] = as.numeric(AccDat$steponset[i] - AccDat$steponset[i-1])
  }
}

# Create variable that contains target instantaneous cadence
AccDat$TargIntStepInterval = lapply(AccDat$cond, function(cond) {
  if(cond == "b1" | cond == "b2") {
    TargIntStepInterval = NA
  } else if (cond == "m100" | cond == "mp100" | cond == "mpb10") {
    TargIntStepInterval = 0.6
  } else if (cond == "m125" | cond == "mp125" | cond == "mpb12") {
    TargIntStepInterval = 0.48
  } else if (cond == "m75" | cond == "mp75" | cond == "mpb75") {
    TargIntStepInterval = 0.8
  } else {
    TargIntStepInterval = NA
  }
})
AccDat$TargIntStepInterval = as.numeric(AccDat$TargIntStepInterval)

# Compute the difference between the instantaneous cadence and the target instantaneous cadence
AccDat$IntStepIntervalDev = AccDat$IntStepInterval - AccDat$TargIntStepInterval
AccDat$Rect_IntStepIntervalDev = abs(AccDat$IntStepIntervalDev)

# Remove trials that are duplicated/recoded per conversation with Simone 4-1-20
AccDat = AccDat %>%
  filter(SubNum != 3 | cond != "m100" | !(trial %in% c(4, 10))) %>%
  filter(SubNum != 3 | cond != "m125" | trial != 10) %>%
  filter(SubNum != 3 | cond != "m75" | trial != 10) %>%
  filter(SubNum != 4 | cond != "m100" | trial != 4) %>%
  filter(SubNum != 7 | cond != "m125" | trial != 10) %>%
  filter(SubNum != 9 | cond != "m100" | trial != 2) %>%
  filter(SubNum != 6 | cond != "mp125" | trial != 5) %>%
  filter(SubNum != 6 | cond != "m100" | trial != 12) %>%
  filter(SubNum != 17 | cond != "m125" | !(is.na(trial))) %>%
  filter(SubNum != 7 | cond != "mpb75" | trial != 12) %>%
  filter(SubNum != 13 | cond != "mpb10" | trial != 12) %>%
  filter(SubNum != 13 | cond != "mp100" | trial != 10 | StepNum != 8) %>%
  filter(SubNum != 13 | cond != "mp100" | trial != 10) %>%
  filter(SubNum != 11 | cond != "b1" | trial != 5)
  

```

## Project Overview

This code imports, analyzes, and reports results evaluating deviations from metronome cues an experiment assessing stepping accuracy between individuals with obesity and healthy-weight controls during a metronome gait task from the Motor Development Lab (Gill).  These data were collected in 2014.  This code was created in March 2020 By Phil Desrochers.

<br>

***

## Assumptions/Definitions

* Forthcoming....

<br>

***

<!-- ## Variable Glossary -->

<!-- * idonset: Onset time of id variable from MacSHAPA. -->
<!-- * idoffset: Offset time of id variable from MacSHAPA. -->
<!-- * latonset: Onset time of latency variable from MacSHAPA. -->
<!-- * latoffset: Offset time of latency variable from MacSHAPA. -->
<!-- * behonset: Onset time of behavioral code variable from MacSHAPA. -->
<!-- * behoffset: Offset time of behavioral code variable from MacSHAPA. -->
<!-- * relonset: Onset time of reliability code variable from MacSHAPA. -->
<!-- * reloffset: Offset time of reliability code variable from MacSHAPA. -->
<!-- * gaitonset: Onset time of gaitdisruptnum variable from MacSHAPA. -->
<!-- * gaitoffset: Offset time of gaitdisruptnum variable from MacSHAPA. -->
<!-- * steponset: Onset time of step variable (with gaitmat data) from MacSHAPA. -->
<!-- * stepoffset: Offset time of step variable (with gaitmat data) from MacSHAPA. -->
<!-- * subj: Subject number for study SXM3 (e.g. S1M3) for adults and SKXM3 (e.g. SK1M3) for children -->
<!-- * cora: Child or adult. -->
<!-- * condord: Condition order - can be A through F. -->
<!-- * gender: Subject's gender - can be m or f. -->
<!-- * tday: Test day. -->
<!-- * bday: Birthday. -->
<!-- * legr: Measurement of right leg in cm. -->
<!-- * legl: Measurement of left leg in cm. -->
<!-- * weight: Measurement of weight in kg. -->
<!-- * bmi: Body mass index weight/height squared ((kg)/m squared)). -->
<!-- * ht: Height in cm. -->
<!-- * injury: Whether subject has had an injury - can by y or n. -->
<!-- * sports: Whether subject has played sports - can be y or n. -->
<!-- * sportsinj: Whether subject has sustained sports injuries - can be y or n. -->
<!-- * whatinj: What subject has injured - letter indicates where injury occurred. -->
<!-- * legdiscrepancy: Whether subject is aware that they have any leg length discrepancy - y or n. -->
<!-- * spinalabn: Whether subject has spinal abnormalities/scoliosis - y or n. -->
<!-- * handedness: Which hand is the subject's dominant hand - r or l. -->
<!-- * footedness: Which hand is the subject's dominant foot - r or l. -->
<!-- * expmetro: Whether subject has had experience walking to a metronome beat - y or n. -->
<!-- * expmusic: Whether subject has had musical or dance experience - y or n. -->
<!-- * condtria: Condition number + trial number. -->
<!-- * cond: Condition - b1 (initial baseline). -->
<!--   * m100 (metronome at 100bpm). -->
<!--   * mp100 (metronome at 100bpm after 2 minutes of practice). -->
<!--   * mpb100 (carryover baseline trials after all trials with metronome at 100bpm). -->
<!--   * m125 (metronome at 125bpm). -->
<!--   * mp125 (metronome at 125bpm after 2 minutes of practice). -->
<!--   * mpb125 (carryover baseline trials after all trials with metronome at 125bpm). -->
<!--   * m75 (metronome at 75bpm). -->
<!--   * mp75 (metronome at 75bpm after 2 minutes of practice). -->
<!--   * mpb75 (carryover baseline trials after all trials with metronome at 75bpm). -->
<!--   * b2 (final baseline). -->
<!-- * lattrip: Whether subject tripped during latency - y or n. -->
<!-- * latfall: Whether subject fell during latency - y or n. -->
<!-- * latstumble: Whether subject stumbled during latency - y or n. -->
<!-- * latlook: Whether subject looked down during latency - y or n. -->
<!-- * lattap: Whether subject tapped during latency - y or n. -->
<!-- * latnod: Whether subject nodded during latency - y or n. -->
<!-- * latcount: Whether subject counted during latency - y or n. -->
<!-- * trip: Whether subject tripped during trial - y or n. -->
<!-- * fall: Whether subject fell during trial - y or n. -->
<!-- * stumble: Whether subject stumbled during trial - y or n. -->
<!-- * stop: Whether subject stopped during trial - y or n. -->
<!-- * crossbackstep: Whether subject cross-stepped or back-stepped during trial - y or n. -->
<!-- * dblstep: Whether subject double-stepped during trial - y or n. -->
<!-- * armsout: Whether subject put arms out to the sides during trial - y or n. -->
<!-- * look: Whether subject looked down during trial - y or n. -->
<!-- * tap: Whether subject tapped during trial - y or n. -->
<!-- * nod: Whether subject nodded during trial - y or n. -->
<!-- * count: Whether subject counted during trial - y or n. -->
<!-- * tripcount: How many times subject tripped during trial (positive integer). -->
<!-- * fallcount: How many times subject fell during trial (positive integer). -->
<!-- * stumblecount: How many times subject stumbled during trial (positive integer). -->
<!-- * stopcount: How many times subject stopped during trial (positive integer). -->
<!-- * crosscount: How many times subject cross or back stepped during trial (positive integer). -->
<!-- * armscount: How many times subject put arms out to the sides during trial (positive integer). -->
<!-- * lookcount: How many times subject looked down during trial (positive integer). -->
<!-- * tapcount: How many times subject tapped during trial (positive integer). -->
<!-- * nodcount: How many times subject nodded during trial (positive integer). -->
<!-- * countcount: How many times subject counted during trial (positive integer). -->
<!-- * distance: Distance in cm walked across the gait carpet. -->
<!-- * velocity: Velocity in cm/s. -->
<!-- * cadence: Cadence. -->
<!-- * heelx: X-coordinate of heel. -->
<!-- * heely: Y-coordinate of heel. -->
<!-- * toex: X-coordinate of toe. -->
<!-- * toey: Y-coordinate of toe. -->
<!-- * firstcontact: First contact time for foot. -->
<!-- * lastcontact: Last contact time for foot. -->
<!-- * lrfoot: Left or right foot (0=left; 1=right). -->
<!-- * stpl: Step length in cm. -->
<!-- * wid: Step width in cm. -->
<!-- * stpt: Step time in seconds. -->
<!-- * swgt: Swing time in seconds. -->
<!-- * stnt: Stance time in seconds. -->
<!-- * sst: Single support time in seconds. -->
<!-- * dst: Double support time in seconds. -->

<!-- <br> -->

<!-- *** -->

## Data Processing Methodology

<br>

### Participant demographics:

Current sample size: n = 15

```{r examine subjet demographics, warning=FALSE, message=FALSE, fig.align="center", fig.width=8, fig.height=5, fig.cap="Figure 1 : Boxplots of Age and BMI, overlayed with individual subject data"}

# Age
AccDat_Age = AccDat %>%
  select(SubNum, testagey) %>%
  group_by(SubNum) %>%
  summarize(Age = mean(testagey))

Age_boxplot = ggplot(data = AccDat_Age, aes(x = 0, y = Age))+
  stat_boxplot(geom = "errorbar", colour = "black", size = 1, width = 0.55)+
  geom_boxplot(fill = '#A3CCAF')+
  geom_dotplot(binaxis = "y", stackdir = "center")+
  theme_bw()+
  theme(axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 14),
        panel.border = element_rect(color = 'black', fill = NA, size = 2),
        panel.grid = element_line(color = "gray"),
        plot.title = element_text(hjust = 0.5, size = 14),
        legend.position = 'none')+
  ggtitle('Age')

# BMI
AccDat_BMI = AccDat %>%
  select(SubNum, realbmi) %>%
  group_by(SubNum) %>%
  summarize(BMI = mean(realbmi))

BMI_boxplot = ggplot(data = AccDat_BMI, aes(x = 0, y = BMI))+
  stat_boxplot(geom = "errorbar", colour = "black", size = 1, width = 0.55)+
  geom_boxplot(fill = '#A3CCAF')+
  geom_dotplot(binaxis = "y", stackdir = "center")+
  theme_bw()+
  theme(axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 14),
        panel.border = element_rect(color = 'black', fill = NA, size = 2),
        panel.grid = element_line(color = "gray"),
        plot.title = element_text(hjust = 0.5, size = 14),
        legend.position = 'none')+
  ggtitle('BMI')

Demo_plots = Age_boxplot + BMI_boxplot
Demo_plots

overweight = count(AccDat_BMI, BMI > 25)

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

**NOTE:** 8 out of 15 participants have overweight BMI

</div>

<br>


### Inter-Step Interval {.tabset .tabset-fade .tabset-pills}

1) Used steponset to compute inter-step time (I called this instantaneous cadence)
2) Computed target instantaneous cadence for each metronome tempo
    * For 75 bpm: (75 bpm / 60 sec)^-1 = 0.8 secs/step
    * For 100 bpm: (100 bpm / 60 sec)^-1 = 0.6 secs/step
    * For 125 bpm: (125 bpm / 60 sec)^-1 = 0.48 sec/step
3) Computed instantaneous deviation by computing difference between instantaneous cadence and target cadence (targ_cadence - inst_cadence)
4) Computed rectified instantaneous deviation by taking absolute value.
5) Averaged both rectified and Raw interstep intervals across steps, trials, and participants within each condition.

<br>

<!-- **Next, I looked to see how many steps each person had in each trial of each condition**. -->

<!-- <br> -->

<!-- #### Steps in each trial in each condition {.tabset .tabset-fade .tabset-pills} -->

<!-- ##### Sub 1 -->

<!-- ```{r Sub1 step table} -->

<!-- AccDat_StepsInTrials = AccDat %>% -->
<!--   select(SubNum, cond, trial, cadence, StepNum, IntStepInterval, IntStepIntervalDev, Rect_IntStepIntervalDev) %>% -->
<!--   filter(cond %in% c("m75", "m100", "m125", "mp75", "mp100", "mp125", "mpb75", "mpb10", "mpb12"))# %>% -->
<!--   #filter(trial <= 10) # subject 6 has a trial #12 for condition m100 - my guess is that the condition should be mp100 -->
<!-- AccDat_StepsInTrials$cond = factor(AccDat_StepsInTrials$cond, levels = c("m75", "m100", "m125", "mp75", "mp100", "mp125", "mpb75", "mpb10", "mpb12")) -->

<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 1) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->

<!-- ``` -->

<!-- ##### Sub 3 -->

<!-- ```{r Sub3 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 3) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 4 -->

<!-- ```{r Sub4 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 4) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 5 -->

<!-- ```{r Sub5 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 5) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 6 -->

<!-- ```{r Sub6 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 6) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 7 -->

<!-- ```{r Sub7 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 7) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 8 -->

<!-- ```{r Sub8 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 8) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 9 -->

<!-- ```{r Sub9 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 9) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 10 -->

<!-- ```{r Sub10 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 10) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 11 -->

<!-- ```{r Sub11 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 11) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 13 -->

<!-- ```{r Sub13 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 13) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 16 -->

<!-- ```{r Sub16 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 16) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 17 -->

<!-- ```{r Sub17 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 17) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 22 -->

<!-- ```{r Sub22 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 22) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- ##### Sub 33 -->

<!-- ```{r Sub33 step table} -->
<!-- subdata = AccDat_StepsInTrials %>% -->
<!--   filter(SubNum == 33) %>% -->
<!--   select(trial, cond, StepNum) %>% -->
<!--   group_by(trial, cond) %>% -->
<!--   summarise(numSteps = n()) %>% -->
<!--   spread(cond, numSteps) -->
<!-- datatable(subdata, rownames = FALSE, filter="top", options = list(pageLength = 12, scrollX=T)) -->
<!-- ``` -->

<!-- #### -->

<!-- <br> -->

#### Raw ISI

<br>

**Descriptive Statistics**

<br>

```{r Raw ISI Descriptives, warning=FALSE, message = FALSE}

AccDat_StepInt_ANOVAdf = AccDat %>%
  select(SubNum, cond, trial, realbmi, cadence, StepNum, IntStepInterval, IntStepIntervalDev, Rect_IntStepIntervalDev) %>%
  filter(cond %in% c("m75", "m100", "m125", "mp75", "mp100", "mp125")) %>%
  filter(StepNum > 1) # %>%
  # group_by(cond, SubNum, trial) %>%
  # summarize_all(funs(mean, std.error), na.rm = TRUE)

AccDat_StepInt_ANOVAdf$cond = factor(AccDat_StepInt_ANOVAdf$cond, levels = c("m75", "m100", "m125", "mp75", "mp100", "mp125"))
AccDat_StepInt_ANOVAdf$BPM = sapply(AccDat_StepInt_ANOVAdf$cond, function(cond) {
  if(cond == "m75" | cond == "mp75" | cond == "mpb75") {
    BPM = 75
  } else if (cond == "m100" | cond == "mp100" | cond == "mpb10") {
    BPM = 100
  } else if (cond == "m125" | cond == "mp125" | cond == "mpb12") {
    BPM = 125
  }
})
AccDat_StepInt_ANOVAdf$Practice = sapply(AccDat_StepInt_ANOVAdf$cond, function(cond) {
  if(cond == "m75" | cond == "m100" | cond == "m125") {
    Practice = "Pre-Practice"
  } else if (cond == "mp75" | cond == "mp100" | cond == "mp125") {
    Practice = "Post-Practice"
  } else
    Practice = NA
})
AccDat_StepInt_ANOVAdf$BPM = factor(AccDat_StepInt_ANOVAdf$BPM)
AccDat_StepInt_ANOVAdf$Practice = factor(AccDat_StepInt_ANOVAdf$Practice, levels = c("Pre-Practice", "Post-Practice"))

AccDat_StepInt_DESC = ezStats(data = AccDat_StepInt_ANOVAdf,
                                    wid = SubNum,
                                    dv = IntStepIntervalDev,  #_mean
                                    within = c(BPM, Practice))
AccDat_StepInt_DESC = AccDat_StepInt_DESC %>%
  mutate(std.error = SD/sqrt(N))
datatable(AccDat_StepInt_DESC, filter="top", options = list(scrollX=T)) %>%
  formatRound(columns = c('Mean', 'SD', 'FLSD', 'std.error'), digits = 5)
```

<br>

**Within-subjects 2 (Practice COndition) x 3 (Pace) rmANOVA**

<br>

```{r raw ISI ANOVA, message=FALSE, warning=FALSE, comment=NA}
AccDat_StepInt_ANOVA = aov_ez(data = AccDat_StepInt_ANOVAdf,
                                    id = 'SubNum',
                                    dv = 'IntStepIntervalDev',  #_mean
                                    within = c('BPM', 'Practice'))
nice(AccDat_StepInt_ANOVA)
```

<br>

**Within subjects contrasts, Tukey corrected**

<br>

Contrasts: Practice at each Pace
```{r raw ISI emmeans Practice at each BPM, warning=FALSE, message = FALSE, comment = NA}
AccDat_StepInt_PracByBPM_emmeans = emmeans(AccDat_StepInt_ANOVA, specs = ~ Practice | BPM)
pairs(AccDat_StepInt_PracByBPM_emmeans)
```

<br>

Contrasts: Pace
```{r raw ISI emmeans Pace, warning=FALSE, message = FALSE, comment = NA}
AccDat_StepInt_BPM_emmeans = emmeans(AccDat_StepInt_ANOVA, specs = ~BPM)
pairs(AccDat_StepInt_BPM_emmeans)
```

<br>

**Bar graph of within subjects contrasts**

<br>

```{r raw ISI barplot, warning=FALSE, message = FALSE, fig.align="center", fig.width=8, fig.height=5, fig.cap="Figure 2 : Raw Inter-step Interval plotted for each pace and practice condition"}
StepInterval_Raw_agg_plot = ggplot(data = AccDat_StepInt_DESC, aes(x = as.factor(BPM), y = Mean, group = Practice, fill = Practice))+
  geom_bar(stat = 'identity', position = "dodge", color = "black")+
  geom_errorbar(aes(ymax = Mean + std.error, ymin = Mean - std.error), size = 1, position = position_dodge(width = 0.9), width = 0.3)+
  labs(fill = "Condition")+
  theme_bw()+
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title = element_text(size = 14),
        panel.border = element_rect(color = 'black', fill = NA, size = 2),
        panel.grid = element_line(color = "gray"),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size = 14),
        plot.title = element_text(hjust = 0.5, size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))+
  ggtitle('Raw Inter-step Interval')
StepInterval_Raw_agg_plot

```

<br>

#### Rectified ISI

<br>

**Descriptive Statistics**

<br>

```{r Rect ISI Descriptives, warning=FALSE, message = FALSE}
AccDat_RectStepInt_DESC = ezStats(data = AccDat_StepInt_ANOVAdf,
                                    wid = SubNum,
                                    dv = Rect_IntStepIntervalDev,  #_mean
                                    within = c(BPM, Practice))
AccDat_RectStepInt_DESC = AccDat_RectStepInt_DESC %>%
  mutate(std.error = SD/sqrt(N))
datatable(AccDat_RectStepInt_DESC, filter="top", options = list(scrollX=T)) %>%
  formatRound(columns = c('Mean', 'SD', 'FLSD', 'std.error'), digits = 5)
```

<br>

**Within-subjects 2 (Practice COndition) x 3 (Pace) rmANOVA**

<br>

```{r Rect ISI ANOVA, message=FALSE, warning=FALSE, comment=NA}
AccDat_RectStepInt_ANOVA = aov_ez(data = AccDat_StepInt_ANOVAdf,
                                    id = 'SubNum',
                                    dv = 'Rect_IntStepIntervalDev',  #_mean
                                    within = c('BPM', 'Practice'))
nice(AccDat_RectStepInt_ANOVA)
```

<br>

**Within subjects contrasts, Tukey corrected**

<br>

Contrasts: Practice at each Pace
```{r Rect ISI emmeans Practice at each BPM, warning=FALSE, message = FALSE, comment = NA}
AccDat_RectStepInt_PracByBPM_emmeans = emmeans(AccDat_RectStepInt_ANOVA, specs = ~ Practice | BPM)
pairs(AccDat_RectStepInt_PracByBPM_emmeans)
```

<br>

Contrasts: Pace
```{r Rect ISI emmeans Pace, warning=FALSE, message = FALSE, comment = NA}
AccDat_RectStepInt_BPM_emmeans = emmeans(AccDat_RectStepInt_ANOVA, specs = ~BPM)
pairs(AccDat_RectStepInt_BPM_emmeans)
```

<br>

**Bar graph of within subjects contrasts**

<br>

```{r Rect ISI barplot, warning=FALSE, message = FALSE, fig.align="center", fig.width=8, fig.height=5, fig.cap="Figure 3 : Rectified Inter-step Interval plotted for each pace and practice condition"}
StepInterval_Rect_agg_plot = ggplot(data = AccDat_RectStepInt_DESC, aes(x = as.factor(BPM), y = Mean, group = Practice, fill = Practice))+
  geom_bar(stat = 'identity', position = "dodge", color = "black")+
  geom_errorbar(aes(ymax = Mean + std.error, ymin = Mean - std.error), size = 1, position = position_dodge(width = 0.9), width = 0.3)+
  labs(fill = "Condition")+
  theme_bw()+
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title = element_text(size = 14),
        panel.border = element_rect(color = 'black', fill = NA, size = 2),
        panel.grid = element_line(color = "gray"),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size = 14),
        plot.title = element_text(hjust = 0.5, size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))+
  ggtitle('Rectified Inter-step Interval')
StepInterval_Rect_agg_plot

```

<br>

###

<br><br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

**CONCLUSION: When examining Raw step interval, no effects of practice are evident.  However, at the fast pace, participants tend to have a longer ISI relative to the target pace, compared to the 75 bpm condition (but not the 100 bpm condition).  No differences exist in rectified ISI.**

**NOTE: I'm pretty sure is not possible to do a fully within subjects ANCOVA (i.e., repeat these analyses covarying for BMI/age/etc).  To accomplish this, we'd probably need to do a linear mixed effects model.**

</div>

<br>

***

### Cadence

1) I computed the difference between the participant's cadence and target cadence (targ_cadence - cadence)
2) Examined cadence deviation between each pace and practice condition.

<br>

```{r Cadence Descriptives , warning=FALSE, message = FALSE, fig.align="center", fig.width=8, fig.height=5, fig.cap="Figure 4: Individual Cadence across trials within conditions"}

AccDat_CadenceDev_ANOVAdf = AccDat %>%
  select(SubNum, cond, trial, cadence, StepNum, IntStepInterval, IntStepIntervalDev, Rect_IntStepIntervalDev) %>%
  filter(cond %in% c("m75", "m100", "m125", "mp75", "mp100", "mp125")) %>%
  filter(StepNum <= 6 & StepNum > 1)# %>%
  # filter(trial <= 10)
AccDat_CadenceDev_ANOVAdf$CadenceTarg = sapply(AccDat_CadenceDev_ANOVAdf$cond, function(cond) {
  if(cond == "m75" | cond == "mp75" | cond == "mpb75") {
    BPM = 75
  } else if (cond == "m100" | cond == "mp100" | cond == "mpb10") {
    BPM = 100
  } else if (cond == "m125" | cond == "mp125" | cond == "mpb12") {
    BPM = 125
  }
})
AccDat_CadenceDev_ANOVAdf$BPM = sapply(AccDat_CadenceDev_ANOVAdf$cond, function(cond) {
  if(cond == "m75" | cond == "mp75" | cond == "mpb75") {
    BPM = 75
  } else if (cond == "m100" | cond == "mp100" | cond == "mpb10") {
    BPM = 100
  } else if (cond == "m125" | cond == "mp125" | cond == "mpb12") {
    BPM = 125
  }
})
AccDat_CadenceDev_ANOVAdf$Practice = sapply(AccDat_CadenceDev_ANOVAdf$cond, function(cond) {
  if(cond == "m75" | cond == "m100" | cond == "m125") {
    Practice = "Pre-Practice"
  } else if (cond == "mp75" | cond == "mp100" | cond == "mp125") {
    Practice = "Post-Practice"
  } else
    Practice = NA
})
AccDat_CadenceDev_ANOVAdf$BPM = factor(AccDat_CadenceDev_ANOVAdf$BPM)
AccDat_CadenceDev_ANOVAdf$Practice = factor(AccDat_CadenceDev_ANOVAdf$Practice, levels = c("Pre-Practice", "Post-Practice"))

AccDat_CadenceDev_ANOVAdf = AccDat_CadenceDev_ANOVAdf %>%
  mutate(CadenceDev = CadenceTarg - cadence)
AccDat_CadenceDev_ANOVAdf$cond = factor(AccDat_CadenceDev_ANOVAdf$cond, levels = c("m75", "m100", "m125", "mp75", "mp100", "mp125", "mpb75", "mpb10", "mpb12"))

AccDat_CadenceDev_DESC = ezStats(data = AccDat_CadenceDev_ANOVAdf,
                                    wid = SubNum,
                                    dv = CadenceDev,  #_mean
                                    within = c(BPM, Practice))
AccDat_CadenceDev_DESC = AccDat_CadenceDev_DESC %>%
  mutate(std.error = SD/sqrt(N))
datatable(AccDat_CadenceDev_DESC, filter="top", options = list(scrollX=T)) %>%
  formatRound(columns = c('Mean', 'SD', 'FLSD', 'std.error'), digits = 5)
```

<br>

**Within-subjects 2 (Practice COndition) x 3 (Pace) rmANOVA**

<br>

```{r Cadence ANOVA, message=FALSE, warning=FALSE, comment=NA}
AccDat_CadenceDev_ANOVA = aov_ez(data = AccDat_CadenceDev_ANOVAdf,
                                    id = 'SubNum',
                                    dv = 'CadenceDev',  #_mean
                                    within = c('BPM', 'Practice'))
nice(AccDat_CadenceDev_ANOVA)
```

<br>

**Within subjects contrasts, Tukey corrected**

<br>

Contrasts: Practice at each Pace
```{r Cadecnce emmeans Practice at each BPM, warning=FALSE, message = FALSE, comment = NA}
AccDat_CadenceDev_PracByBPM_emmeans = emmeans(AccDat_CadenceDev_ANOVA, specs = ~ Practice | BPM)
pairs(AccDat_CadenceDev_PracByBPM_emmeans)
```

<br>

Contrasts: Pace
```{r Cadence emmeans Pace, warning=FALSE, message = FALSE, comment = NA}
AccDat_CadenceDev_BPM_emmeans = emmeans(AccDat_CadenceDev_ANOVA, specs = ~BPM)
pairs(AccDat_CadenceDev_BPM_emmeans)
```

<br>

**Bar graph of within subjects contrasts**

<br>

```{r Cadence barplot, warning=FALSE, message = FALSE, fig.align="center", fig.width=8, fig.height=5, fig.cap="Figure 2 : Raw Inter-step Interval plotted for each pace and practice condition"}
CadenceDev_agg_plot = ggplot(data = AccDat_CadenceDev_DESC, aes(x = as.factor(BPM), y = Mean, group = Practice, fill = Practice))+
  geom_bar(stat = 'identity', position = "dodge", color = "black")+
  geom_errorbar(aes(ymax = Mean + std.error, ymin = Mean - std.error), size = 1, position = position_dodge(width = 0.9), width = 0.3)+
  labs(fill = "Condition")+
  theme_bw()+
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title = element_text(size = 14),
        panel.border = element_rect(color = 'black', fill = NA, size = 2),
        panel.grid = element_line(color = "gray"),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size = 14),
        plot.title = element_text(hjust = 0.5, size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))+
  ggtitle('Raw Inter-step Interval')
CadenceDev_agg_plot

```

<br>

<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

**CONCLUSION: Participants walk at a slower cadence than the target at the 125 BMP time frame**

</div>

<br>

***

## ISI Baseline Comparisons

<br>
 
```{r ISI Baseline Comparison, warning=FALSE, message = FALSE, fig.align="center", fig.width=8, fig.height=5}
AccDat_BL_ttestdf = AccDat %>%
  select(SubNum, cond, trial, cadence, StepNum, IntStepInterval, IntStepIntervalDev, Rect_IntStepIntervalDev) %>%
  filter(cond %in% c("b1", "mpb75", "mpb10", "mpb12")) %>%
  filter(StepNum > 1) %>%
  group_by(SubNum, cond, trial) %>%
  summarize_all(funs(mean(., na.rm = TRUE))) %>%
  ungroup() %>%
  group_by(SubNum, cond) %>%
  summarize_all(funs(mean(., na.rm = TRUE), std.error(., na.rm = TRUE)))
AccDat_BL_ttestdf$cond = factor(AccDat_BL_ttestdf$cond, levels = c("b1", "mpb75", "mpb10", "mpb12"))

B1_v_MPB75_ISI_ttest = t.test(subset(AccDat_BL_ttestdf$IntStepInterval_mean, AccDat_BL_ttestdf$cond == "b1" & AccDat_BL_ttestdf$SubNum != 11), subset(AccDat_BL_ttestdf$IntStepInterval_mean, AccDat_BL_ttestdf$cond == "mpb75" & AccDat_BL_ttestdf$SubNum != 11), paired = TRUE)
B1_v_MPB75_ISI_ttest

B1_v_MPB10_ISI_ttest = t.test(subset(AccDat_BL_ttestdf$IntStepInterval_mean, AccDat_BL_ttestdf$cond == "b1"), subset(AccDat_BL_ttestdf$IntStepInterval_mean, AccDat_BL_ttestdf$cond == "mpb10"), paired = TRUE)
B1_v_MPB10_ISI_ttest

B1_v_MPB12_ISI_ttest = t.test(subset(AccDat_BL_ttestdf$IntStepInterval_mean, AccDat_BL_ttestdf$cond == "b1" & AccDat_BL_ttestdf$SubNum != 13), subset(AccDat_BL_ttestdf$IntStepInterval_mean, AccDat_BL_ttestdf$cond == "mpb12" & AccDat_BL_ttestdf$SubNum != 13), paired = TRUE)
B1_v_MPB12_ISI_ttest

ggplot(data = AccDat_BL_ttestdf, aes(x = cond, y = IntStepInterval_mean))+
  geom_bar(stat = "summary")+
  geom_errorbar(aes(ymin = IntStepInterval_mean - IntStepInterval_std.error, ymax = IntStepInterval_mean - IntStepInterval_std.error), stat = "summary")


B1_v_MPB75_cadence_ttest = t.test(subset(AccDat_BL_ttestdf$cadence_mean, AccDat_BL_ttestdf$cond == "b1" & AccDat_BL_ttestdf$SubNum != 11), subset(AccDat_BL_ttestdf$cadence_mean, AccDat_BL_ttestdf$cond == "mpb75" & AccDat_BL_ttestdf$SubNum != 11), paired = TRUE)
B1_v_MPB75_cadence_ttest

B1_v_MPB10_cadence_ttest = t.test(subset(AccDat_BL_ttestdf$cadence_mean, AccDat_BL_ttestdf$cond == "b1"), subset(AccDat_BL_ttestdf$cadence_mean, AccDat_BL_ttestdf$cond == "mpb10"), paired = TRUE)
B1_v_MPB10_cadence_ttest

B1_v_MPB12_cadence_ttest = t.test(subset(AccDat_BL_ttestdf$cadence_mean, AccDat_BL_ttestdf$cond == "b1" & AccDat_BL_ttestdf$SubNum != 13), subset(AccDat_BL_ttestdf$cadence_mean, AccDat_BL_ttestdf$cond == "mpb12" & AccDat_BL_ttestdf$SubNum != 13), paired = TRUE)
B1_v_MPB12_cadence_ttest

ggplot(data = AccDat_BL_ttestdf, aes(x = cond, y = cadence_mean))+
  geom_bar(stat = "summary")+
  geom_errorbar(aes(ymin = cadence_mean - cadence_std.error, ymax = cadence_mean - cadence_std.error), stat = "summary")

```


## Baseline ISI and Cadence

```{r}

AccDat_StepInt_BL = AccDat %>%
  select(SubNum, cond, trial, cadence, StepNum, IntStepInterval) %>%
  filter(cond %in% c("b1", "b2")) %>%
  filter(StepNum > 1) %>%
  select(cond, cadence, IntStepInterval) %>%
  group_by(cond) %>%
  summarise_all(list(mean, std.error), na.rm = TRUE)



```

***
## Open Questions

<br>

* How to covary BMI?  Do you still think it's beneficial?
* Isn't Inter-step Interval, aggregated across both steps and trials, just a proxy for cadence?  I.e., aren't they saying the exact same things?
* To examine the effects of BMI, we'd have to do a mixed model analysis (not an ANOVA), because I don't think you can use a continuous covariate in a fully within-subjects design.


